\chapter*{Acknowledgment}

I would like to express my deepest gratitude to Professor Dr. Mohammad Hamdaqa from Polytechnique Montréal for
his guidance and support throughout this research journey.
I am immensely thankful for the opportunity he provided, the intellectually stimulating conversations we shared,
and the academic freedom he granted me to explore and develop my research ideas.
His encouragement to engage with other experts significantly enriched this work.

I am profoundly grateful to Dr. Maged Elaasar from NASA Jet Propulsion Laboratory (JPL)
and Professor Dr. Bentley James Oakes from Polytechnique Montréal
for their invaluable insights and constructive feedback.
Their expertise and generosity with their time allowed me to refine my ideas and enhance the overall quality of this
thesis.

My sincere thanks go to Professor Dr. ir. Chris Develder at Ghent University for his administrative support,
detailed feedback, and thorough review of the manuscript.
His commitment to excellence and attention to detail were crucial in ensuring this thesis achieved the desired quality.

I also wish to thank François Goybet for granting me the opportunity to utilize his bachelor project,
which proved invaluable in testing my proof of concept.
I had the pleasure of meeting him through Prof. Dr. Hamdaqa during my exchange in Montréal,
where he was also participating in an exchange program at the time.
His collaboration and generosity in sharing his work were sincerely appreciated.

I am deeply grateful to the Faculty of Engineering and Architecture (FEA) and Ghent University for enabling me to
undertake this research in collaboration with Professor Hamdaqa.
This experience has been a highlight of my academic journey.
I extend my thanks to the International Relations Office of FEA for facilitating my study exchange and ensuring a
rewarding experience, as well as to Polytechnique Montréal for hosting me as an exchange student.

Finally, I want to acknowledge the broader society for the opportunities that allowed me to pursue higher education and
embark on this academic journey.
The privilege to engage in research and study abroad is something I deeply value.

Above all, my heartfelt gratitude goes to my mother.
Her unwavering support, encouragement, and belief in me have been the cornerstone of my achievements.
Her strength continues to inspire me every day.

